---
title: "Final Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
fao <- read_csv("FAO.csv") 
head(fao)
dim(fao) 
```

```{r}
sub_fao <- subset(unique(fao[, c("Area Abbreviation", "Area")]), select = c("Area Abbreviation", "Area"))

# print the two columns next to each other
print(sub_fao)
```

```{r}
#install "units" package
library(units)

```

```{r}
sub_food_fao <- subset(unique(fao[, c("Item Code", "Item")]), select = c("Item Code", "Item"))
# print the two columns next to each other
head(sub_food_fao)

# use recode() to collapse levels into categories
fao1 <- fao
fao1$Item %>% discard(is.null)

fao1$Item <- recode(fao1$Item, "Potatoes and products" = "Vegetable", "Vegetables, Other" = "Vegetable",
                         "Vegetables" = "Vegetable", "Cassava and products" = "Vegetable", "Sweet potatoes" = "Vegetable",
                         "Beans" = "Vegetable", "Onions" = "Vegetable", "Yams" = "Vegetable")

fao1$Item <- recode(fao1$Item, "Soyabean Oil" = "Oil", "Groundnut Oil" = "Oil",
                     "Sunflowerseed Oil" = "Oil", "Rape and Mustard Oil" = "Oil", "Cottonseed Oil" = "Oil",
                   "Palm Oil" = "Oil", "Sesameseed Oil" = "Oil", "Olive Oil" = "Oil", "Oilcrops Oil, Other" = "Oil",
                   "Oilcrops" = "Oil", "Vegetable Oils" = "Oil", "Oilcrops, Other" = "Oil", "Maize Germ Oil" = "Oil",
                   "Fish, Body Oil" = "Oil", "Fish, Liver Oil" = "Oil", "Coconut Oil" = "Oil", "Palmkernel Oil" = "Oil",
                   "Ricebran Oil" = "Oil")

fao1$Item <- recode(fao1$Item, "Coffee and products" = "Beverage", "Tea (including mate)" = "Beverage", "Wine" = "Beverage",
                        "Beer" = "Beverage", "Beverages, Alcoholic" = "Beverage", "Alcoholic Beverages" = "Beverage",
                        "Beverages, Fermented" = "Beverage")

fao1$Item <- recode(fao1$Item, "Bovine Meat" = "Meat", "Mutton & Goat Meat" = "Meat", "Poultry Meat" = "Meat",
                    "Offals, Edible" = "Meat", "Meat, Other" = "Meat", "Freshwater Fish" = "Meat", "Meat" = "Meat",
                   "Fish, Seafood" = "Meat", "Pigmeat" = "Meat", "Demersal Fish" = "Meat",
                    "Pelagic Fish" = "Meat", "Marine Fish, Other" = "Meat", "Crustaceans" = "Meat", "Cephalopods" = "Meat",
                    "Molluscs, Other" = "Meat", "Meat, Aquatic Mammals" = "Meat")

fao1$Item <- recode(fao1$Item, "Butter, Ghee" = "AnimalProducts", "Cream" = "AnimalProducts",
                        "Fats, Animals, Raw" = "AnimalProducts", "Eggs" = "AnimalProducts",
                        "Milk - Excluding Butter" = "AnimalProducts", "Eggs" = "AnimalProducts",
                        "Milk - Excluding Butter" = "AnimalProducts", "Animal fats" = "AnimalProducts",
                        "Offals" = "AnimalProducts")

fao1$Item <- recode(fao1$Item, "Wheat and products" = "Grain", "Rice (Milled Equivalent)" = "Grain",
                       "Barley and products" = "Grain", "Maize and products" = "Grain", "Millet and products" = "Grain",
                       "Cereals, Other" = "Grain", "Pulses, Other and products" = "Grain", "Pulses" = "Grain",
                     "Rye and products" = "Grain", "Oat" = "Grain", "Sorghum and products" = "Grain", "Peas" = "Grain",
                     "Soyabeans" = "Grain", "Palm kernels" = "Grain", "Sunflower seed" = "Grain", "Cottonseed" = "Grain")

fao1$Item <- recode(fao1$Item, "Nuts and products" = "Fruits", "Coconuts - Incl Copra" = "Fruits", "Olives (including preserved)" = "Fruits", "Tomatoes and products" = "Fruits", "Oranges, Mandarines" = "Fruits", "Citrus, Other" = "Fruits",
                      "Bananas" = "Fruits", "Apples and products" = "Fruits", "Pineapples and products" = "Fruits",
                      "Dates" = "Fruits", "Grapes and products (excl wine)" = "Fruits", "Fruits, Other" = "Fruits",
                      "Treenuts" = "Fruits", "Fruits - Excluding Wine" = "Fruits", "Groundnuts (Shelled Eq)" = "Fruits",
                      "Lemons, Limes and products" = "Fruits", "Grapefruit and products" = "Fruits",
                      "Plantains" = "Fruits",  "Pimento" = "Fruits")

fao1$Item <- recode(fao1$Item, "Sesame seed" = "Other", "Cocoa Beans and products" = "Other", "Pepper" = "Other",
                     "Spices, Other" = "Other", "Sugar cane" = "Other", "Sugar beet" = "Other","Sugar (Raw Equivalent)" = "Other", "Sweeteners, Other" = "Other", "Honey" = "Other", "Infant food" = "Other", "Cereals - Excluding Beer" = "Other",
                     "Starchy Roots" = "Other", "Sugar Crops" = "Other", "Sugar & Sweeteners" = "Other",
                     "Stimulants" = "Other", "Spices" = "Other", "Miscellaneous" = "Other", "Roots, Other" = "Other",
                     "Rape and Mustardseed" = "Other", "Aquatic Animals, Others" = "Other", "Aquatic Plants" = "Other",
                     "Aquatic Products, Other" = "Other", "Cloves" = "Other", "Sugar non-centrifugal" = "Other")

sub_food_fao1 <- subset(unique(fao1[, c("Item Code", "Item")]), select = c("Item Code", "Item"))
# print the two columns next to each other
head(sub_food_fao1)

```

Time series plot
```{r}
  ggplot(fao1, aes(x = c(Y1961,2013), y = Element, group=Item, color = Item)) + geom_line()
```
map plot
```{r}
head(fao1)

ggplot(fao1, aes( x=longitude, y=latitude, color=Area)) + geom_point()



library(ggplot2)
library(maps)
library(mapdata)

# create a data frame with latitude and longitude coordinates
fao1 <- fao1(longitude, latitude)

# create a world map using the maps package
world_map <- world_map("world")

# plot the map and add the points for the locations data frame
ggplot() +
  geom_polygon(data = world_map, aes(x = longitude, y = latitude, group = group), fill = "gray") +
  geom_point(data = fao1, aes(x = longitude, y = latitude), color = "red") +
  labs(title = "Map with latitude and longitude coordinates")
```

























```{r}
#bar plot
map <- ggplot(fao1, aes(Item, ..count..)) + geom_bar(aes(fill = Area) + facet_grid(cols = vars(Year %in% c(Y1961, Y1971, Y1981, Y1991, Y2001, Y2011))), position ="dodge")
map
```


 
```{r}
library(sf)
library(mapview)
mapspots <- st_as_sf(fao1, coords = c("latitude", "longitude"),  crs = 4326)
mapspots
mapview(mapspots, map.types = "Stamen.Toner")
```


```{r}
fao1 %>% ggplot(aes(x=longitude, y=latitude)) + geom_point()
```











=======
Tess Kern, Leah Greiner, Maheeka Devarakondra
https://www.kaggle.com/datasets/dorbicycle/world-foodfeed-production?resource=download
We went to kaggle to find a data set focusing on agriculture. We are both interested in getting jobs in the agriculture field, so we wanted to look at something agricultural related. Our data set looks at the growing population and where food is going. There are two main uses for the food/grain grown; food and feed. There are 63 columns in this data set, and over 21,000 rows. The data is represented since 1961 to 2013 and considers countries like Czechoslovakia and South Sudan and whether they are still a country or are a newly formed country since 1961. If the country is no longer, then it has been removed from the dataset. If the country has been newly formed since 1961, it is in the data set, but just has limited data. 
Research Question: Is there an increase in a certain type of crop going towards feed or food.
#data cleaning
We need to determine what each country's abbreviation is. We also need to update the columns to the correct year. We need to convert the unit into the rest of the columns because each country has a different unit. We need to sort the items into categories such as vegetables, fruits, meats, and other to make the data easier to evaluate.
>>>>>>> 7d6c7ccb4a98c53b6ebf0a470b1c74d498bc5489
