---
title: "Final Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
We chose a data set that deals with crops grown around the world and some of their end uses.
```{r}
library(tidyverse)
library(ggplot2)
fao <- read_csv("FAO.csv") 
head(fao)
dim(fao) 
```

```{r}
sub_fao <- subset(unique(fao[, c("Area Abbreviation", "Area")]), select = c("Area Abbreviation", "Area"))

# printing the two columns next to each other
print(sub_fao)
#This was how we determined each countries abbreviation. We were able to sort through the data without any repeats. 
```

```{r}
sub_food_fao <- subset(unique(fao[, c("Item Code", "Item")]), select = c("Item Code", "Item"))
# printing the two columns next to each other
head(sub_food_fao)

# using recode() to collapse levels into categories
fao1 <- fao
fao1$Item %>% discard(is.null)

fao1$Item <- recode(fao1$Item, "Potatoes and products" = "Vegetable", "Vegetables, Other" = "Vegetable",
                         "Vegetables" = "Vegetable", "Cassava and products" = "Vegetable", "Sweet potatoes" = "Vegetable",
                         "Beans" = "Vegetable", "Onions" = "Vegetable", "Yams" = "Vegetable")

fao1$Item <- recode(fao1$Item, "Soyabean Oil" = "Oil", "Groundnut Oil" = "Oil",
                     "Sunflowerseed Oil" = "Oil", "Rape and Mustard Oil" = "Oil", "Cottonseed Oil" = "Oil",
                   "Palm Oil" = "Oil", "Sesameseed Oil" = "Oil", "Olive Oil" = "Oil", "Oilcrops Oil, Other" = "Oil",
                   "Oilcrops" = "Oil", "Vegetable Oils" = "Oil", "Oilcrops, Other" = "Oil", "Maize Germ Oil" = "Oil",
                   "Fish, Body Oil" = "Oil", "Fish, Liver Oil" = "Oil", "Coconut Oil" = "Oil", "Palmkernel Oil" = "Oil",
                   "Ricebran Oil" = "Oil")

fao1$Item <- recode(fao1$Item, "Coffee and products" = "Beverage", "Tea (including mate)" = "Beverage", "Wine" = "Beverage",
                        "Beer" = "Beverage", "Beverages, Alcoholic" = "Beverage", "Alcoholic Beverages" = "Beverage",
                        "Beverages, Fermented" = "Beverage")

fao1$Item <- recode(fao1$Item, "Bovine Meat" = "Meat", "Mutton & Goat Meat" = "Meat", "Poultry Meat" = "Meat",
                    "Offals, Edible" = "Meat", "Meat, Other" = "Meat", "Freshwater Fish" = "Meat", "Meat" = "Meat",
                   "Fish, Seafood" = "Meat", "Pigmeat" = "Meat", "Demersal Fish" = "Meat",
                    "Pelagic Fish" = "Meat", "Marine Fish, Other" = "Meat", "Crustaceans" = "Meat", "Cephalopods" = "Meat",
                    "Molluscs, Other" = "Meat", "Meat, Aquatic Mammals" = "Meat")

fao1$Item <- recode(fao1$Item, "Butter, Ghee" = "AnimalProducts", "Cream" = "AnimalProducts",
                        "Fats, Animals, Raw" = "AnimalProducts", "Eggs" = "AnimalProducts",
                        "Milk - Excluding Butter" = "AnimalProducts", "Eggs" = "AnimalProducts",
                        "Milk - Excluding Butter" = "AnimalProducts", "Animal fats" = "AnimalProducts",
                        "Offals" = "AnimalProducts")

fao1$Item <- recode(fao1$Item, "Wheat and products" = "Grain", "Rice (Milled Equivalent)" = "Grain",
                       "Barley and products" = "Grain", "Maize and products" = "Grain", "Millet and products" = "Grain",
                       "Cereals, Other" = "Grain", "Pulses, Other and products" = "Grain", "Pulses" = "Grain",
                     "Rye and products" = "Grain", "Oat" = "Grain", "Sorghum and products" = "Grain", "Peas" = "Grain",
                     "Soyabeans" = "Grain", "Palm kernels" = "Grain", "Sunflower seed" = "Grain", "Cottonseed" = "Grain", "Oats" = "Grain")

fao1$Item <- recode(fao1$Item, "Nuts and products" = "Fruits", "Coconuts - Incl Copra" = "Fruits", "Olives (including preserved)" = "Fruits", "Tomatoes and products" = "Fruits", "Oranges, Mandarines" = "Fruits", "Citrus, Other" = "Fruits",
                      "Bananas" = "Fruits", "Apples and products" = "Fruits", "Pineapples and products" = "Fruits",
                      "Dates" = "Fruits", "Grapes and products (excl wine)" = "Fruits", "Fruits, Other" = "Fruits",
                      "Treenuts" = "Fruits", "Fruits - Excluding Wine" = "Fruits", "Groundnuts (Shelled Eq)" = "Fruits",
                      "Lemons, Limes and products" = "Fruits", "Grapefruit and products" = "Fruits",
                      "Plantains" = "Fruits",  "Pimento" = "Fruits")

fao1$Item <- recode(fao1$Item, "Sesame seed" = "Other", "Cocoa Beans and products" = "Other", "Pepper" = "Other",
                     "Spices, Other" = "Other", "Sugar cane" = "Other", "Sugar beet" = "Other","Sugar (Raw Equivalent)" = "Other", "Sweeteners, Other" = "Other", "Honey" = "Other", "Infant food" = "Other", "Cereals - Excluding Beer" = "Other",
                     "Starchy Roots" = "Other", "Sugar Crops" = "Other", "Sugar & Sweeteners" = "Other",
                     "Stimulants" = "Other", "Spices" = "Other", "Miscellaneous" = "Other", "Roots, Other" = "Other",
                     "Rape and Mustardseed" = "Other", "Aquatic Animals, Others" = "Other", "Aquatic Plants" = "Other",
                     "Aquatic Products, Other" = "Other", "Cloves" = "Other", "Sugar non-centrifugal" = "Other")

sub_food_fao1 <- subset(unique(fao1[, c("Item Code", "Item")]), select = c("Item Code", "Item"))
# printing the two columns next to each other
head(sub_food_fao1)

```
There wasn't much cleaning that needed to be done. The most strenuous task was recoding the data, so we weren't dealing with too many variables. This took up quite a bit of time, but we made it through. We ran into a problem with some of the recoded data, and had to remove any NULL values. We also had issues making sure everything was put under the column "Item" when we recoded. For example: "fao1$Item <- recode(fao1$meat," did not work, so we changed 'meat' to 'Item' and that solved our issue. 


```{r, warning=FALSE}
#Bar Chart
ggplot(fao1, aes(x = Area, fill = Item)) +
    geom_bar(stat = "count") 
```


```{r}
#bar chart
ggplot(fao1, aes(x = Item)) +
    geom_bar(stat = "count") 
```

```{r}
#bar chart
countries <- c("USA", "AZE", "SRB", "EGY", "LBR")

fao2 <- subset(fao1, `Area Abbreviation` %in% countries)

ggplot(fao2, aes(x = Area, fill = Item)) +
    geom_bar(stat = "count") 

```


```{r}
library(ggmosaic)

ggplot(data = fao2)  + 
  geom_mosaic(aes(x = product(Area), fill=Item, weight=1))
```

```{r}
#map plot
fao1 %>% ggplot(aes(x=longitude, y=latitude)) + geom_point()
fao1
#We couldn't get the world map to work as a backdrop for the plot. But, this graph does a justice of what we found. It is representing the latitude and longitude spots of all the countries represented in the data set. You can see the outline of various countries in Africa as well as a large population of countries located in Europe.
```

 
```{r}
#histogram

countries <- c("LBR", "USA", "SRB", "ISL", "IND")
 
fao3 <- subset(fao1, `Area Abbreviation` %in% countries)

ggplot(fao3, aes(Area)) + geom_bar(aes(fill = Element), position = "dodge")
#This hisogram show the differences between 5 countries selected at random in their feed and food exports. This is total across all years.
```

```{r}
#scatter
ggplot(fao1, aes(x = Y2000, y = Unit)) + geom_point() + geom_jitter() + labs(x = "", y = "Value", title = "Year 2000 crop production across all countries")
#This scatterplot shows the relationship between 199 and 2000 values. You can see a strong linear relationship. This means that values between 1999 and 2000 are similar, as if the value was low in 1999, it was low in 2000 as well.
```
```{r}
agg_data <- aggregate(data$Area, by = list(data$Area), FUN = mean)

# Rename columns
colnames(agg_data) <- c("Country", "Mean_Variable")


# Fix the subset command using square brackets
subset_data <- data[Area == "USA", ]



# Print aggregated data
agg_data

boxplot(Mean_Variable ~ Country, data = agg_data, 
        xlab = "Country", ylab = "Mean Variable")
ggplot(data = fao3, aes(x = "United States of America", y = Item)) + geom_boxplot() + coord_flip()
```

```{r}
head(fao)
ggplot(fao2, aes(x = Area, y = Item)) + geom_line() 
```
 
 













=======
Tess Kern, Leah Greiner, Maheeka Devarakondra
https://www.kaggle.com/datasets/dorbicycle/world-foodfeed-production?resource=download
We went to kaggle to find a data set focusing on agriculture. We are both interested in getting jobs in the agriculture field, so we wanted to look at something agricultural related. Our data set looks at the growing population and where food is going. There are two main uses for the food/grain grown; food and feed. There are 63 columns in this data set, and over 21,000 rows. The data is represented since 1961 to 2013 and considers countries like Czechoslovakia and South Sudan and whether they are still a country or are a newly formed country since 1961. If the country is no longer, then it has been removed from the dataset. If the country has been newly formed since 1961, it is in the data set, but just has limited data. 
Research Question: Is there an increase in a certain type of crop going towards feed or food.
#data cleaning
We need to determine what each country's abbreviation is. We also need to update the columns to the correct year. We need to convert the unit into the rest of the columns because each country has a different unit. We need to sort the items into categories such as vegetables, fruits, meats, and other to make the data easier to evaluate.
>>>>>>> 7d6c7ccb4a98c53b6ebf0a470b1c74d498bc5489
